Biden Projected to Win in the 2024 US Presidential Election

### 2024-03-15 paper/paper.qmd
Data section added

The paper is further organised into three sections: Data, Model, Results, and Discussion. In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed. We will explain the variables of interest in the datasets for the analysis. The Model section describes the model being used for the analyis. The Results section will then highlight and discuss the trends and associations found during the analysis. Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses of our analysis.

@@

##Introduction to the Data:
We will use the most recent American Community Survey, the 2022 Sample, as the post-stratification dataset
to represent the voter population and adjust the weight of the survey data. The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected. ACS 2022 dataset consists of 3,373,378 observations.

Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic,
socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population.

Since the ACS 2022 sample is weighted, the data are guaranteed to accurately reflect the demographic makeup of the country's population. Weighting improves the reliability and accuracy of statistical analyses and conclusions generated from the dataset by facilitating more precise extrapolation of findings from the sample to the larger population.

The ACS 2022 sample is noteworthy for its geographic structure, where the Public Use Microdata Area (PUMA) is the
smallest recognizable geographic unit.
Geographic areas that do not cross state boundaries and have a population of at least 100,000 are known as PUMAs.
This physical separation makes it easier to analyze regional differences and trends in socioeconomic and demographic traits, enabling scholars to investigate the subtleties of voting patterns and electoral dynamics throughout the nation.

All things considered, the ACS 2022 sample is a useful tool for statistical analysis and empirical study, providing insights into the complex aspects of American culture. In the sections that follow, we explore the variables that make up the dataset and consider how important they are to comprehending voting patterns and electoral results for the 2024 election using this 20022 survey data.

## Introduction to the Variables
These are the key variables utilized in our statistical analysis of voting behavior during the 2022 survey.
These variables are drawn from the American Community Survey (ACS) 2022 sample and encompass demographic,
socio-economic, and geographic characteristics that may influence individuals' voting preferences.

Age: Age is a representation of each candidate's actual age as of the election. A key demographic factor that might affect voting patterns, generational dynamics, and political opinions is age.

Income: The pre-tax amount of the respondents' families' entire annual income from all sources is shown as income. As a stand-in for socioeconomic position and financial security, income has the power to influence people's political choices, attitudes, and worldviews.

Education: Based on the highest year of school or degree earned, respondents' educational attainment is indicated by their education. knowledge electoral dynamics requires a knowledge of the important variable of educational attainment, which is directly connected to political ideology, civic participation, and voter turnout.

State: The respondents' state of residence provides information about where their housing units are located and makes it possible to analyze local politics and voting trends. The political climate, the priorities of policies, and the demographic makeup of a state can all affect how people vote and how elections turn out.

Sex: The report of sex indicates a person's gender. Political beliefs and actions are greatly influenced by gender dynamics, which may have an impact on voting patterns and election results.

>>>>

The paper is further organised into three sections: Data, Model, Results, and Discussion. In the Data section, we discuss how the datasets to be used for the analysis were obtained and pre-processed. We will explain the variables of interest in the datasets for the analysis. The Model section describes the model being used for the analysis. The Results section will then highlight and discuss the trends and associations found during the analysis. Lastly, the Discussion section will talk about some interesting trends found in Results in depth, link it to the real world and also highlight the weaknesses of our analysis.

@@

For this analysis, we have used two datasets. The datasets were cleaned and analysed using the statistical programming software R [@citeR] along with the help tidyverse [@citeTidyverse], knitr [@citeKnitr], ggplot2 [@citeGgplot], here [@citeHere], ADD MORE

## Survey Data

Our survey data is from the Ipsos Poll [Roper #31120637]. It was collated from a web-based survey conducted on November 3 - 4, 2023 using KnowledgePanel. The survey is conducted by Ipsos and the sponsor is ABC News. This poll is based on a nationally representative probability sample of 949 adults age 18 or older with oversamples among Black and Hispanic respondents. The poll is a collection of demographic information and political preferences.

A person becomes an entry in this dataset if they were 18 years or older on November 3-4, 2023, an American national and participated in this poll. They are recruited through a scientifically developed addressed-based sampling methodology using the latest Delivery Sequence File of the USPS – a database with full coverage of all delivery points in the US. Households invited to join the panel are randomly selected from all available households in the US. People in the sampled households are invited to join and participate in the panel. Those selected who do not already have internet access are provided a tablet and internet connection at no cost to the panel member. Those who join the panel and who are selected to participate in a survey are sent a unique password-protected log-in used to complete surveys online. Panelists receive a unique login to the survey and are only able to complete it one time. No reminder emails were sent for this study.

The results from this dataset are used for our model. Among the range of variables available, we selected age, income, education, state and sex as our independent variables.

- Age is the age group the panelist falls into. The poll had the ages of the panelists at the time of the poll which we then classified into one of the four age groups: 18-29, 30-44, 45-59, 60+

- Income is the annual household income reported by the panelist. They were given the option of seven categories: Less than $10,000, $10,000 to $24,999, $25,000 to $49,999, $50,000 to $74,999, $75,000 to $99,999, $100,000 to $149,999 and $150,000 or more.

- Education is the highest level of education attained by the panelist. They were given the option between: Less than high school, High school, Some college and Bachelor's degree or higher. We then classified these into three categories:  Less than high school, High school,  More than high school

- State is the state of residence of the panelist.

- Sex is the sex reported by the panelist and it is either male or female

For the variable we are predicting, we created a binary variable which is 1 if the panelist favors Biden and 0 if the panelist favors Trump. There was no question in the poll which asked this direct question so we used “Overall, do you have a favorable or unfavorable impression of...Joe Biden?”. We acknowledge that not favoring Joe Biden does not necessarily imply favoring Donald Trump, but since they are the two major potential candidates, we classify the panelists who chose “Unfavorable” as supporting Trump, i.e not supporting Biden.

## Post-stratification Data

Our post-stratification data is from the most recent American Community Survey (ACS), the 2022 Sample. We use this to represent the voter population and adjust the weight of the survey data. The population for this sample was the entire US population, categorized into households as subgroups, and the responses of the sampled subgroups were collected. ACS 2022 dataset consists of 3,373,378 observations. Using a 1-in-100 national random sample, the ACS 2022 sample provides a thorough overview of the demographic, socioeconomic, and geographic features of the US population. Most importantly, this dataset includes people who live in both home and group quarters, offering a comprehensive view of the American population. A person becomes an entry in this dataset if they are a part of the sampled subgroups and have responded to the survey.

There are a large number of different features available and the variables we chose are age, income, education, state and sex to match with the variables from our survey. These variables were cleaned and the dataset was preprocessed so the values and categories match with the survey dataset.
